Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-07-15 doc/bcspaper/article.tex
Started changing things to appease reviewers

Since the advent of the `Digital Age', the amount of information available to
researchers has been increasing drastically, relevant material is becoming
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as AGRICOLA (\url{http://agricola.nal.usda.gov/}),
the Cochrane Library(\url{http://www.thecochranelibrary.com/}) and Textpresso
(\url{http://www.textpresso.org/}) index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain.

The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv and PloSOne that try to incorporate wide ranges of
papers from as many disciplines as possible. The traits of these systems often
complement those of their domain-specific counterparts; they provide a
comprehensive collection of literature but insufficient filtering and indexing
capabilities, usually based on title and abstract.

However, the document title is just one of the crucial parts of a scientific
paper's structure.  Liakata \emph{et al.} (2012) describe a system for
automatically processing and classifying sentences in a research paper
according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis},
{\em Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.  SAPIENTA
(\url{http://www.sapientaproject.com}) is a publicly available machine learning
application which can automatically annotate all sentences in a scientific
paper with their CoreSC labels. It was trained using a corpus of physical
chemistry and biochemistry research papers whose sentences were manually
annotated using the CoreSC\cite{LIAKATA10.644} scheme.  An intelligent 
information retrieval system can use this data to provide better filtering and
search capabilities for researchers.  Partridge implements such context-aware
keyword search, by allowing researchers to search for papers where a term
appears in sentences with a specific CoreSC label (e.g only in {\em Method}
sentences). This can be used to greatly improve both the precision with which
researchers are able to perform searches for scientific literature and the
accuracy of those searches in terms of relevance to the reader.

@@

% in conclusions add how this can give recommendations for Plos and other journals 
The papers were downloaded from PLoS using {\em plosget.py}, a short
python script that uses the PLoS RESTful search API. The number of
papers downloaded per paper type category was as follows: 200 Essay,
99 Correspondence, 107 Case Study, 200 Perspective, 74 Viewpoint, 93
Opinion, 312 Review, 200 Research.  These formed a corpus of 1285 %1459-174 after removing synopsis. Should double check the numbers add up
papers.

@@

A confusion matrix for the paper types in given in Table
\ref{tab:confusion}. Research papers are usually predicted to belong
to the Research class, and are sometimes confused with Case
Study. However, Case Study papers are often predicted to be Research
or Review. The ratio between Case Study, Research and Review papers is
~1:2:3 so the outcome is not entirely surprising in terms of the data
size effect. We expected Research and Case Study to share similar
paper structures but perhaps the fact that Case Study is equally
confused with Research and Review suggests two distinct types of Case
Study.  We expected the classes Essay, Opinion, Perspective and
Viewpoint to be confused, as the four labels all indicate a paper
containing an author's personal thoughts on an issue, rather than
experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except
Research. Viewpoint is a small class and mostly confused with Opinion
and Perspective. From this it would seem that the paper types
\{Opinion, Perspective, Viewpoint\} are very similar and should be
grouped together into one category. Indeed when we trained the model
on a super class consisting of the previous three categories,
performance was improved overall with the following F-measures:
{\bf OpinionSuper: 59, Research: 67.6, Review: 63.7, Essay: 44.9,
Correspondence: 46 and Case Study 24.8}.

@@

\vspace{-15pt}
\section{Conclusion}
\label{sec:4}

To summarise, we have demonstrated that paper type can largely be
predicted from the distribution of the automatically obtained CoreSC
sentence labels in the full text of a paper. We have described some of
the particular CoreSC features that determine a paper type
(e.g. Correspondence characterised by very little Background, Opinions
characterised by low percentage of Experiment) and discussed which of
the paper types are not easily separable in this way. We recommended
for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall
classification performance. It is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making
process to be complex, but agrees with our general understanding of
paper types. The Partridge online system now allows refinement of
paper search using CoreSC and paper type, both of which can be
intelligently determined using machine learning methods.

>>>>

Since the advent of the `Digital Age', the amount of information available to
researchers has been increasing drastically, relevant material is becoming
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as AGRICOLA \footnote{\url{http://agricola.nal.usda.gov/}},
the Cochrane Library\footnote{\url{http://www.thecochranelibrary.com/}} and Textpresso
\footnote{\url{http://www.textpresso.org/}} index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain.

The value of these systems to their users is often restricted by the small
proportion of available literature that they index, forcing researchers to use
multiple, domain specific, search engines for their queries.  In contrast,
there are also a number of interdisciplinary indexing systems and online
journals such as arXiv \footnote{\url{http://arxiv.org}} and PloSOne
\footnote{\url{http://plosone.org}} that try to incorporate wide ranges of
papers from as many disciplines as possible. The traits of these systems often
complement those of their domain-specific counterparts; they provide a
comprehensive collection of literature but insufficient filtering and indexing
capabilities, usually based on title and abstract.

However, the document title is just one of the crucial parts of a scientific
paper's structure.  Liakata \emph{et al.} (2012) describe a system for
automatically processing and classifying sentences in a research paper
according to the core scientific concept (CoreSC) that they
describe\cite{Liakata2012}.  There are 11 CoreSCs, including {\em Hypothesis},
{\em Goal}, {\it Background}, {\em Method}, {\em Result} and {\em Conclusion}.
CoreSC labels can be allocated to all sentences in a scientific paper in order
to identify which scientific concept each sentence encapsulates.  SAPIENTA
\footnote{\url{http://www.sapientaproject.com}} is a publicly available machine learning
application which can automatically annotate all sentences in a scientific
paper with their CoreSC labels. It was trained using a corpus of physical
chemistry and biochemistry research papers whose sentences were manually
annotated using the CoreSC\cite{LIAKATA10.644} scheme.  An intelligent 
information retrieval system can use this data to provide better filtering and
search capabilities for researchers.  Partridge implements such context-aware
keyword search, by allowing researchers to search for papers where a term
appears in sentences with a specific CoreSC label (e.g only in {\em Method}
sentences). This can be used to greatly improve both the precision with which
researchers are able to perform searches for scientific literature and the
accuracy of those searches in terms of relevance to the reader.

@@

% in conclusions add how this can give recommendations for Plos and other
% journals 
A script called {\em plosget.py} was written, in order to download the
papers via the PloS RESTful search API \footnote{\url{http://api.plos.org/solr/faq/}}. The number of papers downloaded per
paper type category was as follows: 200 Essay, 99 Correspondence, 107 Case
Study, 200 Perspective, 74 Viewpoint, 93 Opinion, 312 Review, 200 Research.
These formed a corpus of 1285 %1459-174 after removing synopsis. Should double
check the numbers add up papers.

@@

A confusion matrix for the paper types is given in Table
\ref{tab:confusion}. Research papers are usually predicted to belong
to the Research class, and are sometimes confused with Case
Study. However, Case Study papers are often predicted to be Research
or Review. The ratio between Case Study, Research and Review papers is
~1:2:3 so the outcome is not entirely surprising in terms of the data
size effect. We expected Research and Case Study to share similar
paper structures but perhaps the fact that Case Study is equally
confused with Research and Review suggests two distinct types of Case
Study.  We expected the classes Essay, Opinion, Perspective and
Viewpoint to be confused, as the four labels all indicate a paper
containing an author's personal thoughts on an issue, rather than
experiment-driven science. Opinion is confused with Perspective and
Viewpoint. Perspective is confused with almost all classes, except
Research. Viewpoint is a small class and mostly confused with Opinion
and Perspective. From this it would seem that the paper types
\{Opinion, Perspective, Viewpoint\} are very similar and should be
grouped together into one category. Indeed when we trained the model
on a super class consisting of the previous three categories,
performance was improved overall with the following F-measures:
{\bf OpinionSuper: 59, Research: 67.6, Review: 63.7, Essay: 44.9,
Correspondence: 46 and Case Study 24.8}.

@@

\vspace{-15pt}
\section{Summary and Conclusions}
\label{sec:4}

To summarise, we have demonstrated that paper type can largely be predicted
from the distribution of the automatically obtained CoreSC sentence labels in
the full text of a paper. We have described some of the particular CoreSC
features that determine a paper type (e.g. Correspondence characterised by very
little Background, Opinions characterised by low percentage of Experiment) and
discussed which of the paper types are not easily separable in this way. We
recommended for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall classification performance. It
is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making process to be complex,
but agrees with our general understanding of paper types. The Partridge system
allows refinement of paper search using CoreSC and paper type, both of which
can be intelligently determined using machine learning methods.
